Searching inscriptions through the EAGLE Portal

### 2015-12-18 source/XX/EAGLE_2016_submission_XX.tex
Added title to XX

%%%%%%%%
%TITLE:%
%%%%%%%%

\title{}
\author{claudio prandoni}
\date{2015-11-09}
\begin{document}
[Warning: Draw object ignored]Abstract

\ \ \ \ \ \ This paper describes the functionalities and the technical infrastructure of the EAGLE Portal, the main
gateway into the EAGLE’s massive epigraphic database. The portal can be accessed both by a human through a web browser
and by external applications through a set of APIs. It is possible to perform full-text searches using a simple
interface, or to launch more advanced queries, including the possibility to upload an image and search for inscriptions
that are similar to the provided one. The seven controlled multilingual vocabularies that were created to help aligning
the multilingual metadata of the inscriptions from the different content providers, have been also integrated in the
search engine as well as all the translations the are available on the EAGLE MediaWiki. All this makes of the EAGLE
Portal a very useful tool for the epigraphic community but also for single users and enthusiasts willing to contribute
to the research in this field.

Keywords

Search engine, API, image recognition, controlled vocabulary, multilingual translations, advanced search, citizen
participation.

1.1. Introduction

The EAGLE Portal is the place where the content provided by the epigraphers’ community is aggregated and stored and
where it is made accessible to the users. It is composed by:

\begin{itemize}
\item The Metadata Aggregation System (in the following referred also as Aggregator), which is where all the content is
stored and indexed.
\item A Graphical User Interface (GUI), which is the interface through which users can browse the content and interact
with it.
\end{itemize}
The Aggregator relies on the D-NET software toolkit, developed within the DRIVER project and extended within the
OpenAIRE, OpenAIREplus\footnote{http://www.openaire.eu/}, EFG, EFG1914\footnote{http://www.europeanfilmgateway.eu/it}
and HOPE\footnote{http://www.peoplesheritage.eu/} projects. D-NET is an open source, general-purpose software conceived
to enable the realization and operation of Aggregative Data Infrastructures (ADIs) and to facilitate their evolution
and maintenance over time. D-NET implements a service-oriented framework based on standards, namely Web Services with
SOAP and REST APIs, where ADIs can be constructed in a LEGO-like approach, by selecting, customizing, and properly
combining D-NET services as building blocks. The resulting ADIs are systems, which can be customized, further extended
(e.g. by the integration of new services), and scaled (e.g. storage and index replicas can be maintained and deployed
on remote nodes to tackle multiple concurrent accesses or very-large data size) at run-time [Amato et al., 2013].

The GUI exposes all the content stored in the Aggregator providing to the users the following functionalities [Prandoni
et al., 2014].

\begin{itemize}
\item Search and browse the rich set of data made available by EAGLE content providers by using either a free text
search or a more advanced interface, including faceted browsing through the integration of the EAGLE controlled
vocabularies.
\item Similarity search through the image recognition algorithm that have been integrated in the EAGLE Portal.
\item Access to all the peer-reviewed translations of the epigraphic texts, in several European languages that have been
produced and integrated in the EAGLE MediaWiki.
\item Export to the user own PC the EpiDoc document of an object for further analysis and processing.
\item Annotate and save relevant information in a user Personal Space (e.g. records of inscriptions, search results,
queries), including content saved while using the Flagship Mobile Application.
\item Access to the Flagship Storytelling Application to browse the existing epigraphy-related narratives and to create
a new story.
\end{itemize}
The ingestion and curation of data is performed through a separate dedicated interface made available by the Aggregator.
This interface includes a series of functionalities to support data ingestion and storage as well as for importing,
indexing, enriching and managing the harvested metadata [Amato et al., 2015].

1.2. The EAGLE Collections

Classical Greek and Latin culture is at the very foundation of modern European identity. From philosophy to
architecture, geometry to law, a variety of contemporary subjects and disciplines have their roots in the classical
world. Only a small fraction of the total production of Greco-Roman texts has survived to the present day, leaving wide
gaps in the historiographical record of an epoch that is immensely relevant to our modern day lives.

@@

Arachne\footnote{ http://arachne.uni-koeln.de } is the central object-database of the German Archaeological Institute
(DAI). Since 2001, Arachne has been integrating negative archives of ancient sculpture that went beyond the specialised
documentation retained in Cologne itself, such as the negatives of ancient sculpture of the German Archaeological
Institute in Rome and the historic glass plate negative collections. Besides this larger project, many additional
activities are going on different levels, for example the online preparations for the »Corpus der Antiken
Sarkophagreliefs«.

@@

Epigraphic Database Bari (EDB)\footnote{ http://www.edb.uniba.it } is a project specialized in epigraphic documents of
Christian patronage (third to eighth centuries, AD), including those contained in the Inscriptiones Christianae Vrbis
Romae, nova series, voll. I-X, in civitate Vaticana 1922-1992 (= ICVR), and those edited in other bibliographical seats
and/or not contained in the ICVR. Currently the inscriptions present in the EDB (counting those already online and
those awaiting definitive approval) amount to 40,000 items ca, though this number is obviously increasing continually.

The task of the Epigraphic Database Heidelberg (EDH)\footnote{ http://edh-www.adw.uni-heidelberg.de } is the systematic
entry of ancient Latin and bilingual (usually Latin and Greek) inscriptions into a complex database. The Epigraphic
Text Database is the heart of EDH and contains 65,000 inscriptions at present. Almost all of the records present texts,
which have already either been edited in the monumental Inscription corpora or published, revised and discussed in
thousands of scholarly articles.

Epigraphic Database Rome (EDR)\footnote{ http://edr-edr.it } was launched in 1999 as an experimental project aimed at
creating a unified database for ancient epigraphy. EDR is part of the international federation of Epigraphic Databases
called Electronic Archive of Greek and Latin Epigraphy. The federation’s purpose is to collect all published Greek and
Latin inscriptions up to the 7th century A.D. considering their best existing editions, also enclosing, when possible,
a number of additional important data and/or images. As part of the federation, EDR aims at collecting the whole
epigraphy of Rome and of the Italian peninsula including Sardinia and Sicily, with the exception of Christian
inscriptions (under EDB jurisdiction).

Hispania Epigraphica Online\footnote{ http://eda-bea.es} was created in 2002 when an EU grant enabled a joint research
project between the Archivo Epigráfico de Hispania and Mag. K. Schaller, who was developing computing applications for
archaeological purposes. The focus of the collection is the rich epigraphic patrimony of Portugal and Spain, mainly
written in Latin, but with some small pockets of Greek, Semitic and Iberian inscriptions.

Petrae database\footnote{ http://petrae.tge-adonis.fr} is a system for the recording of Latin and Greek inscriptions
developed at the Institut Ausonius, which collects epigraphic texts from the various regions in which its researchers
and collaborators are active. Each record produced has the text of an inscription in both an uppercase and lowercase
version, accompanied by metadata on all aspects of the monument, including support, fragments, epigraphic fields and
text elements (dating, paleography, critical apparatus, translation, notes).

The Last Statues of Antiquity\footnote{ http://laststatues.classics.ox.ac.uk } is a project funded in 2009 by the ‘Arts
and Humanities Research Council’ in Oxford. The project collects and analyses all the evidence for new, newly
dedicated, or newly re-worked statuary in the period circa 284–650. The two main results of the project are a major
database, with over 2600 individual entries, and a book, published in late 2012, discussing in print the entire
phenomenon of the late-antique statue habit.

The picture database VBI ERAT LUPA\footnote{ http://www.ubi-erat-lupa.org } contains stone monuments (sculptures,
reliefs, inscriptions, architectural pieces etc.). The project’s scope ranges from prehistoric stone monuments to
around the time of Justinian (500 AD). Due to its inception in Vienna, most project data at the moment is from the mid-
and south-eastern European region.

Finally, thanks to the great effort in the networking activities, many other partners joined the EAGLE Best Practice
Network and made available their collections that are currently being uploaded in the EAGLE Portal\footnote{ For a full
list of the EAGLE Associated Partners see http://www.eagle-network.eu/about/partners/ }.

1.3. The EAGLE inscription search engine

The EAGLE Inscriptions Search Engine is accessible through the main horizontal navigation bar of the EAGLE
Portal\footnote{ http://www.eagle-network.eu/basic-search/ }. It represents the core functionality of the portal,
through which the entry of keywords and phrases produce matches from EAGLE’s massive epigraphic database.

The EAGLE Portal makes available a “simple search”, an “advanced search” where the user can specify the values of a
number of fields in order to make a more accurate search, and an “image search” where the user can upload an image and
search for inscriptions that look similar to the provided one [Prandoni et al., 2014].

The objects that a user can search in the EAGLE Portal belong to three different categories, in accordance with the
EAGLE conceptual model [Sicila et al., 2015]:

\begin{itemize}
\item “Artefacts”, which contain all the information that is somehow related to the physical carrier of the inscription.
\item “Texts”, which contain all the information that is textual in nature.
\item “Images”, which contain all the information that is visual in nature.
\end{itemize}

\bigskip

1.3.1. How to search inscriptions

\bigskip

The simple search user interface is very straightforward. The text entered in the query box is used to make a full text
search in all the fields of all the EAGLE objects in the category determined when making the query. By default, the
search will be done on all the objects in the “Artefact” category. Two more tabs on the result list (labelled “Texts”
and “Images”) allow the user to perform the same query choosing a different category.

Using the advanced search interface, a user can specify values for a number of fields, in order to have more accurate
results: findspot, bibliography, text of the inscription, type of inscription, decoration, object type, material, type
of writing, state of preservation (Fig. 1).

In the fields having a controlled vocabulary, the user is allowed to enter only values coming from the vocabulary. For
this purpose, those fields have a “drop-down menu” which displays all the defined values for that field in the
“preferred label”, i.e. the text string that has been indicated in the vocabulary as the preferred one for display,
regardless of the language in which the string is defined.

Finally, by choosing the Image Search option, a user can exploit the image recognition algorithm that has been
integrated in the EAGLE Portal and that allows to search for images that are similar to the one a user has uploaded,
ranked in decreasing order of similarity (Fig. 2).

\bigskip

 \includegraphics[width=10.874cm,height=12.885cm]{EAGLE2016submissionXX-img001.png}

Fig. 1. Search for inscriptions in the EAGLE Portal

\bigskip

Content Based Image Retrieval (CBIR) is becoming an effective way for searching digital libraries, as the amount of
available image data is constantly increasing. CBIR applications are increasingly becoming useful in accessing cultural
heritage information, as a complement to metadata based search. In fact, in some cases metadata associated with images
do not describe the content with sufficient details to satisfy the user queries, or metadata are completely missing.
For example, images containing reproductions of works of art contain a lot of implicit information that is not
generally captured in manually or automatically generated metadata [Amato et al., 2013].

@@

\bigskip

 \includegraphics[width=10.848cm,height=10.451cm]{EAGLE2016submissionXX-img002.png}

Fig. 2. Image search in EAGLE

\bigskip

@@

\bigskip

 \includegraphics[width=10.874cm,height=12.197cm]{EAGLE2016submissionXX-img003.png}

Fig. 3. Results list - Artefacts

\bigskip

1.3.2. Filtering and browsing results

\bigskip

After having submitted a query, the user is presented with a paginated list of results containing some basic information
(Fig. 3).

@@

Regardless of the type of query performed by the user and of the category in which the query was done, clicking on one
of the items in the result page will display a summary of all the information available for that item, with links to
get further details (Fig. 4).

\bigskip

 \includegraphics[width=10.874cm,height=8.864cm]{EAGLE2016submissionXX-img004.png}

Fig. 4. Object details

\bigskip

@@

Duplicates identification has been performed through the Trismegistos platform\footnote{ http://www.trismegistos.org/ }.
By standardizing publication references, collection information, material, provenance and date, the overlap of the
various databases has been mapped to identify records describing the same inscription. Afterwards a unique numeric
identifier, the Trismegistos ID, has been assigned to each document and spread across the partner databases for
inclusion in the metadata record to be uploaded in the EAGLE Portal.

If more than one instance of an object is available, the other instances appear on the summary page as clickable “tabs”,
each tab being labelled with the local ID of the other instances of the same object.

The seven controlled multilingual vocabularies that have been created (Type of Inscription, Object Type, Material,
Execution, Decoration, State of Preservation, Dating Criteria) to help aligning the multilingual metadata of the
inscriptions from the different content providers, have been also integrated in the EAGLE search engine. Every time
that a term which belongs to one of the vocabularies is displayed, it is automatically linked to the specific page
describing the term, where further connections, synonyms and translations can be explored. The EAGLE vocabularies are
accessible also as Linked Open Data\footnote{ http://www.eagle-network.eu/resources/vocabularies/}.

Finally, all the peer-reviewed translations in several European languages that are available, have been integrated in
the EAGLE Portal too through a RESTful API native of Wikibase, the MediaWiki extension used for this purpose\footnote{
http://www.eagle-network.eu/wiki/api.php }. For each inscription it is possible to view the existing translation, to
request a new one and to contribute by connecting to the EAGLE MediaWiki\footnote{ http://www.eagle-network.eu/wiki }.

\bigskip

\ 1.3.3. Annotating and saving objects

\bigskip

A registered user, during a “web” or a “mobile” session, has the capability of annotating and saving (some of) the
information that is being provided by the EAGLE system.

@@

1.4. Conclusion

@@

In this framework, the CIVIC EPISTEMOLOGIES\footnote{ http://www.civic-epistemologies.eu } project developed a Roadmap
for the use of e-Infrastructures to support the participation of European citizens in CH practices and humanities
research, where such engagement has a twofold benefit for culture: to be enriched by the citizens’ contributions and to
become more widely used and exploited (also, for example, with the participation of creative industries) [Fresa \&
Justrell, 2015]. The future development of the EAGLE initiative will refer to the CIVIC EPISTEMOLOGIES Roadmap for
investigating how to put in place more advanced services supporting citizen science in epigraphic research.

The EAGLE Portal is only an example of application that makes use of the APIs provided by the Aggregator to access the
EAGLE collections. Other examples are the Flagship Mobile Application\footnote{
http://www.eagle-network.eu/resources/flagship-mobile-app/ }, that allows to access the inscriptions’ database through
a mobile device and to fully exploit the image recognition features integrated in the EAGLE Portal, and the Flagship
Storytelling Application\footnote{ http://www.eagle-network.eu/stories/ }, that allows a user to create stories
starting from the content available in EAGLE, Europeana and many other data sources.

Other applications may be created based on the APIs provided by the Aggregator, which uses a SOLR-base search engine
through which it is possible to search and browse the full set of content\footnote{ For further information see
http://wiki.apache.org/solr/\#Search\_and\_Indexing }. Furthermore, APIs are provided to register a new user, login and
perform an image-based search using the EAGLE image recognition algorithm.

>>>>

%%%%%%%%
%TITLE:%
%%%%%%%%
\title{Searching inscriptions through the EAGLE Portal}

%%%%%%%%
%AUTHORS:%
%%%%%%%%
\author[PROMOTER]{Claudio Prandoni\corref{first}}
% put the label for the address (like FENWAY here) in the square brackets; change the addresses below to match the labelled authors
% mark the corresponding author with the \corref command and fill out the complete address in the \cortext command below

\author[PROMOTER]{Antonella Fresa}
\author[Gogate]{Nicola Alfarano}
\author[ISTI]{Giuseppe Amato}
\author[ISTI]{Franco Zoppi}
\author[ISTI]{Andrea Mannocci}
\author[HEI]{Pietro Liuzzo}
\author[DAI]{Francesco Mambrini}
\author[ROME]{Silvia Orlandi}
\author[ROME]{Raffaella Santucci}
%Promoter Srl1, Gogate Srl2, CNR-ISTI3, Heidelberg University4, German Archeological Institute5, Sapienza University of Rome6

%\author[ENEA]{Francesco Biccari}

%Additional information about corresponding author, addresses and affiliations:
\cortext[first]{Corresponding author. Email:  prandoni@promoter.it}

\address[PROMOTER]{Promoter Srl, Via Boccioni 2, 56037 Peccioli, Italy}
\address[Googate]{Gogate Srl, Via Malasoma 24, 56121 Ospedaletto, Italy}
\address[ISTI]{CNR-ISTI, Via G. Moruzzi 1, 56124 Pisa, Italy }
\address[HEI]{Ruprecht-Karls-Universität Heidelberg, Marstallstraße 6, 69117, Heidelberg (DE)}
%\address[ENEA]{ENEA, Casaccia Research Center, via anguillarese 301, 00123 Roma, Italy}
\address[DAI]{Deutsches Archäologisches Institut, 69-71 Podbielskiallee, 14195 Berlin, Germany}
\address[ROME]{La Sapienza University of Rome, Piazzale Aldo Moro 5, 00185 Roma }

%%%%%%%%%%%%
%PAPER CONTENT %
%start your paper here%
%%%%%%%%%%%%

\begin{document}

\maketitle

\begin{abstract}
This paper describes the functionalities and the technical infrastructure of the EAGLE Portal, the main
gateway into the EAGLE’s massive epigraphic database. The portal can be accessed both by a human through a web browser
and by external applications through a set of APIs. It is possible to perform full-text searches using a simple
interface, or to launch more advanced queries, including the possibility to upload an image and search for inscriptions
that are similar to the provided one. The seven controlled multilingual vocabularies that were created to help aligning
the multilingual metadata of the inscriptions from the different content providers, have been also integrated in the
search engine as well as all the translations the are available on the EAGLE MediaWiki. All this makes of the EAGLE
Portal a very useful tool for the epigraphic community but also for single users and enthusiasts willing to contribute
to the research in this field.
\end{abstract}

\keywords{Search engine, API, image recognition, controlled vocabulary, multilingual translations, advanced search, citizen
participation}

\section{Introduction}

\noindent The EAGLE Portal is the place where the content provided by the epigraphers' community is aggregated and stored and
where it is made accessible to the users. It is composed by:

\begin{itemize}
\item The Metadata Aggregation System (in the following referred also as Aggregator), which is where all the content is
stored and indexed.
\item A Graphical User Interface (GUI), which is the interface through which users can browse the content and interact
with it.
\end{itemize}
The Aggregator relies on the D-NET software toolkit, developed within the DRIVER project and extended within the
OpenAIRE, OpenAIREplus\footnote{\url{http://www.openaire.eu/}}, EFG, EFG1914\footnote{\url{http://www.europeanfilmgateway.eu/it}}
and HOPE\footnote{\url{http://www.peoplesheritage.eu/}} projects. D-NET is an open source, general-purpose software conceived
to enable the realization and operation of Aggregative Data Infrastructures (ADIs) and to facilitate their evolution
and maintenance over time. D-NET implements a service-oriented framework based on standards, namely Web Services with
SOAP and REST APIs, where ADIs can be constructed in a LEGO-like approach, by selecting, customizing, and properly
combining D-NET services as building blocks. The resulting ADIs are systems, which can be customized, further extended
(e.g. by the integration of new services), and scaled (e.g. storage and index replicas can be maintained and deployed
on remote nodes to tackle multiple concurrent accesses or very-large data size) at run-time \citep{amato_aim_2013}.

The GUI exposes all the content stored in the Aggregator providing to the users the following functionalities \citep{prandoni_eagle_2014}.

\begin{itemize}
\item Search and browse the rich set of data made available by EAGLE content providers by using either a free text
search or a more advanced interface, including faceted browsing through the integration of the EAGLE controlled
vocabularies.
\item Similarity search through the image recognition algorithm that have been integrated in the EAGLE Portal.
\item Access to all the peer-reviewed translations of the epigraphic texts, in several European languages that have been
produced and integrated in the EAGLE MediaWiki.
\item Export to the user own PC the EpiDoc document of an object for further analysis and processing.
\item Annotate and save relevant information in a user Personal Space (e.g. records of inscriptions, search results,
queries), including content saved while using the Flagship Mobile Application.
\item Access to the Flagship Storytelling Application to browse the existing epigraphy-related narratives and to create
a new story.
\end{itemize}
The ingestion and curation of data is performed through a separate dedicated interface made available by the Aggregator.
This interface includes a series of functionalities to support data ingestion and storage as well as for importing,
indexing, enriching and managing the harvested metadata \citep{amato_second_2015}.

\section{The EAGLE Collections}

\noindent Classical Greek and Latin culture is at the very foundation of modern European identity. From philosophy to
architecture, geometry to law, a variety of contemporary subjects and disciplines have their roots in the classical
world. Only a small fraction of the total production of Greco-Roman texts has survived to the present day, leaving wide
gaps in the historiographical record of an epoch that is immensely relevant to our modern day lives.

@@

Arachne\footnote{\url{http://arachne.uni-koeln.de}} is the central object-database of the German Archaeological Institute
(DAI). Since 2001, Arachne has been integrating negative archives of ancient sculpture that went beyond the specialised
documentation retained in Cologne itself, such as the negatives of ancient sculpture of the German Archaeological
Institute in Rome and the historic glass plate negative collections. Besides this larger project, many additional
activities are going on different levels, for example the online preparations for the »Corpus der Antiken
Sarkophagreliefs«.

@@

Epigraphic Database Bari (EDB)\footnote{\url{http://www.edb.uniba.it}} is a project specialized in epigraphic documents of
Christian patronage (third to eighth centuries, AD), including those contained in the \emph{Inscriptiones Christianae Vrbis
Romae, nova series}, voll. I-X, in civitate Vaticana 1922-1992 (= ICVR), and those edited in other bibliographical seats
and/or not contained in the ICVR. Currently the inscriptions present in the EDB (counting those already online and
those awaiting definitive approval) amount to 40,000 items ca, though this number is obviously increasing continually.

The task of the Epigraphic Database Heidelberg (EDH)\footnote{\url{http://edh-www.adw.uni-heidelberg.de}} is the systematic
entry of ancient Latin and bilingual (usually Latin and Greek) inscriptions into a complex database. The Epigraphic
Text Database is the heart of EDH and contains 65,000 inscriptions at present. Almost all of the records present texts,
which have already either been edited in the monumental Inscription corpora or published, revised and discussed in
thousands of scholarly articles.

Epigraphic Database Rome (EDR)\footnote{\url{http://edr-edr.it}} was launched in 1999 as an experimental project aimed at
creating a unified database for ancient epigraphy. EDR is part of the international federation of Epigraphic Databases
called Electronic Archive of Greek and Latin Epigraphy. The federation’s purpose is to collect all published Greek and
Latin inscriptions up to the 7th century A.D. considering their best existing editions, also enclosing, when possible,
a number of additional important data and/or images. As part of the federation, EDR aims at collecting the whole
epigraphy of Rome and of the Italian peninsula including Sardinia and Sicily, with the exception of Christian
inscriptions (under EDB jurisdiction).

Hispania Epigraphica Online\footnote{\url{http://eda-bea.es}} was created in 2002 when an EU grant enabled a joint research
project between the Archivo Epigráfico de Hispania and Mag. K. Schaller, who was developing computing applications for
archaeological purposes. The focus of the collection is the rich epigraphic patrimony of Portugal and Spain, mainly
written in Latin, but with some small pockets of Greek, Semitic and Iberian inscriptions.

Petrae database\footnote{\url{http://petrae.tge-adonis.fr}} is a system for the recording of Latin and Greek inscriptions
developed at the Institut Ausonius, which collects epigraphic texts from the various regions in which its researchers
and collaborators are active. Each record produced has the text of an inscription in both an uppercase and lowercase
version, accompanied by metadata on all aspects of the monument, including support, fragments, epigraphic fields and
text elements (dating, paleography, critical apparatus, translation, notes).

The Last Statues of Antiquity\footnote{\url{http://laststatues.classics.ox.ac.uk}} is a project funded in 2009 by the ‘Arts
and Humanities Research Council’ in Oxford. The project collects and analyses all the evidence for new, newly
dedicated, or newly re-worked statuary in the period \textit{circa} 284–650. The two main results of the project are a major
database, with over 2600 individual entries, and a book, published in late 2012, discussing in print the entire
phenomenon of the late-antique statue habit.

The picture database VBI ERAT LUPA\footnote{\url{http://www.ubi-erat-lupa.org}} contains stone monuments (sculptures,
reliefs, inscriptions, architectural pieces etc.). The project’s scope ranges from prehistoric stone monuments to
around the time of Justinian (500 AD). Due to its inception in Vienna, most project data at the moment is from the mid-
and south-eastern European region.

Finally, thanks to the great effort in the networking activities, many other partners joined the EAGLE Best Practice
Network and made available their collections that are currently being uploaded in the EAGLE Portal\footnote{ For a full
list of the EAGLE Associated Partners see \url{http://www.eagle-network.eu/about/partners/}}.

\section{The EAGLE inscription search engine}

\noindent The EAGLE Inscriptions Search Engine is accessible through the main horizontal navigation bar of the EAGLE
Portal\footnote{\url{http://www.eagle-network.eu/basic-search/}}. It represents the core functionality of the portal,
through which the entry of keywords and phrases produce matches from EAGLE’s massive epigraphic database.

The EAGLE Portal makes available a ``simple search'', an ``advanced search'' where the user can specify the values of a
number of fields in order to make a more accurate search, and an ``image search'' where the user can upload an image and
search for inscriptions that look similar to the provided one \citep{prandoni_eagle_2014}.

The objects that a user can search in the EAGLE Portal belong to three different categories, in accordance with the
EAGLE conceptual model \citep{sicilia_eagle_2015}:

\begin{itemize}
\item ``Artefacts'', which contain all the information that is somehow related to the physical carrier of the inscription.
\item ``Texts'', which contain all the information that is textual in nature.
\item ``Images'', which contain all the information that is visual in nature.
\end{itemize}

\subsection{How to search inscriptions}

\noindent The simple search user interface is very straightforward. The text entered in the query box is used to make a full text
search in all the fields of all the EAGLE objects in the category determined when making the query. By default, the
search will be done on all the objects in the ``Artefact'' category. Two more tabs on the result list (labelled ``Texts''
and ``Images'') allow the user to perform the same query choosing a different category.

Using the advanced search interface, a user can specify values for a number of fields, in order to have more accurate
results: findspot, bibliography, text of the inscription, type of inscription, decoration, object type, material, type
of writing, state of preservation (Fig. \ref{fig:figure1}).

In the fields having a controlled vocabulary, the user is allowed to enter only values coming from the vocabulary. For
this purpose, those fields have a ``drop-down menu'' which displays all the defined values for that field in the
``preferred label'', i.e. the text string that has been indicated in the vocabulary as the preferred one for display,
regardless of the language in which the string is defined.

Finally, by choosing the Image Search option, a user can exploit the image recognition algorithm that has been
integrated in the EAGLE Portal and that allows to search for images that are similar to the one a user has uploaded,
ranked in decreasing order of similarity (Fig. \ref{figure2}).

\begin{figure}[!bp]
\centering
 \includegraphics[width=10.874cm,height=12.885cm]{EAGLE2016submissionXX-img001.png} 
\caption{Search for inscriptions in the EAGLE Portal}
\label{fig:figure1}
\end{figure}

Content Based Image Retrieval (CBIR) is becoming an effective way for searching digital libraries, as the amount of
available image data is constantly increasing. CBIR applications are increasingly becoming useful in accessing cultural
heritage information, as a complement to metadata based search. In fact, in some cases metadata associated with images
do not describe the content with sufficient details to satisfy the user queries, or metadata are completely missing.
For example, images containing reproductions of works of art contain a lot of implicit information that is not
generally captured in manually or automatically generated metadata \citep{amato_aim_2013}.

@@

\begin{figure}[!bp]
\centering
 \includegraphics[width=10.848cm,height=10.451cm]{EAGLE2016submissionXX-img002.png} 
\caption{Image search in EAGLE}
\label{figure2}
\end{figure}

@@

\begin{figure}[!bp]
\centering
 \includegraphics[width=10.874cm,height=12.197cm]{EAGLE2016submissionXX-img003.png} 
\caption{Results list - Artefacts}
\label{figure3}
\end{figure}

\subsection{Filtering and browsing results}

\noindent After having submitted a query, the user is presented with a paginated list of results containing some basic information (Fig. \ref{figure3}).

@@

Regardless of the type of query performed by the user and of the category in which the query was done, clicking on one
of the items in the result page will display a summary of all the information available for that item, with links to
get further details (Fig. \ref{figure4}).

\begin{figure}[!bp]
\centering
\includegraphics[width=10.874cm,height=8.864cm]{EAGLE2016submissionXX-img004.png}
\caption{Object details}
\label{figure4}
\end{figure}

@@

Duplicates identification has been performed through the Trismegistos platform\footnote{\url{http://www.trismegistos.org/}}.
By standardizing publication references, collection information, material, provenance and date, the overlap of the
various databases has been mapped to identify records describing the same inscription. Afterwards a unique numeric
identifier, the Trismegistos ID, has been assigned to each document and spread across the partner databases for
inclusion in the metadata record to be uploaded in the EAGLE Portal.

If more than one instance of an object is available, the other instances appear on the summary page as clickable ``tabs'',
each tab being labelled with the local ID of the other instances of the same object.

The seven controlled multilingual vocabularies that have been created (Type of Inscription, Object Type, Material,
Execution, Decoration, State of Preservation, Dating Criteria) to help aligning the multilingual metadata of the
inscriptions from the different content providers, have been also integrated in the EAGLE search engine. Every time
that a term which belongs to one of the vocabularies is displayed, it is automatically linked to the specific page
describing the term, where further connections, synonyms and translations can be explored. The EAGLE vocabularies are
accessible also as Linked Open Data\footnote{\url{http://www.eagle-network.eu/resources/vocabularies/}}.

Finally, all the peer-reviewed translations in several European languages that are available, have been integrated in
the EAGLE Portal too through a RESTful API native of Wikibase, the MediaWiki extension used for this purpose\footnote{\url{http://www.eagle-network.eu/wiki/api.php}}. For each inscription it is possible to view the existing translation, to
request a new one and to contribute by connecting to the EAGLE MediaWiki\footnote{\url{http://www.eagle-network.eu/wiki}}.

\subsection{Annotating and saving objects}

A registered user, during a ``web'' or a ``mobile'' session, has the capability of annotating and saving (some of) the
information that is being provided by the EAGLE system.

@@

\section{Conclusion}

@@

In this framework, the CIVIC EPISTEMOLOGIES\footnote{\url{http://www.civic-epistemologies.eu}} project developed a Roadmap
for the use of e-Infrastructures to support the participation of European citizens in CH practices and humanities
research, where such engagement has a twofold benefit for culture: to be enriched by the citizens’ contributions and to
become more widely used and exploited (also, for example, with the participation of creative industries) \citep{fresa_roadmap_2015}. The future development of the EAGLE initiative will refer to the CIVIC EPISTEMOLOGIES Roadmap for
investigating how to put in place more advanced services supporting citizen science in epigraphic research.

The EAGLE Portal is only an example of application that makes use of the APIs provided by the Aggregator to access the
EAGLE collections. Other examples are the Flagship Mobile Application\footnote{\url{http://www.eagle-network.eu/resources/flagship-mobile-app/}}, that allows to access the inscriptions’ database through
a mobile device and to fully exploit the image recognition features integrated in the EAGLE Portal, and the Flagship
Storytelling Application\footnote{\url{http://www.eagle-network.eu/stories/}}, that allows a user to create stories
starting from the content available in EAGLE, Europeana and many other data sources.

Other applications may be created based on the APIs provided by the Aggregator, which uses a SOLR-base search engine
through which it is possible to search and browse the full set of content\footnote{ For further information see
\url{http://wiki.apache.org/solr/\#Search\_and\_Indexing}}. Furthermore, APIs are provided to register a new user, login and
perform an image-based search using the EAGLE image recognition algorithm.
